### 2021-08-21 manuscript/manuscript.Rmd
Add manuscript Rmd plus appendix and bib

  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links. The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")`). ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

@@

keywords          : "clinical trials, registration, reporting, meta-research"
wordcount         : "3551"

bibliography      : ["references.bib", "r-references.bib"]
# appendix          : "appendix.Rmd"

numbersections    : yes
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

always_allow_html: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
ggplot2::theme_set(ggplot2::theme_minimal())

library(dplyr)
library(tidyr)
library(stringr)
library(glue)
library(purrr)
library(here)
library(readr)

library(ggplot2)
library(DiagrammeR)
library(gtsummary)
library(gt)
library(papaja)
# library(knitr)
# library(kableExtra)

library(rentrez)
library(roadoi)
theme_gtsummary_compact()

# Get data
trials_all <- read_rds(here("data", "processed", "trials.rds"))
pubmed_all <- read_rds(here("data", "processed", "pubmed-main.rds"))

# Prepare references
r_refs("r-references.bib")
r_citation <- cite_r() # file = "r-references.bib", footnote = TRUE, pkgs, withhold #r_citations$r and r_citations$pkgs

# Source analyses
source(here("R", "functions", "screening.R"))
# # source(here("R", "functions", "report-results.R"))
source(here("R", "analysis", "01_screen-trials.R"))
# # source(here("R", "analysis", "02_flowchart-trials.R"))
source(here("R", "analysis", "03_tbl-trials.R"))
source(here("R", "analysis", "04_link-reg-pub.R"))
# # source(here("R", "analysis", "05_upset-link-reg-pub.R"))
source(here("R", "analysis", "06_plot-change-over-time.R"))
source(here("R", "analysis", "trn-regex-analysis.R"))
source(here("R", "analysis", "cross-registrations.R"))
source(here("R", "analysis", "logistic-regression.R"))
source(here("R", "analysis", "si-if-abs-ft.R"))
source(here("R", "analysis", "reg-auto-link.R"))
source(here("R", "analysis", "tbl-stakeholders.R"))
# # fs::dir_ls(here("R", "analysis")) %>% 
# #   purrr::walk(source)
```

```{r abstract-vars, cache = TRUE}
abs_n_trials_link_full <- report_pct_num_denom_trials(n_trials_link_full)
abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
abs_n_trials_link_trn_ft <- "1131, 60% of trials"
  # Change to gtsummary once updated: https://stackoverflow.com/questions/68502977/gtsummaryinline-text-for-stat-0-with-pattern
  # inline_text(tbl_link_by_reg, has_iv_trn_ft_pdf, stat_0, pattern = "{n}, {p}% of trials")
  # report_n_pct_trials(n_trials_link_trn_ft)

abs_n_trials_ctgov_link_reg_trn_si_derived <-
  inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%")
```

@@

Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses. The validity of health evidence synthesis is jeopardized if relevant trial data are left out to do inadequate or incomplete trial metadata, or inability to find trial results through automated means. Incomplete evidence threatens healthcare and the lives and welfare of human volunteers. Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b]. Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->. Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].

@@

The inclusion of TRNs in publication and bibliographic metadata especially enhances machine readability and discoverability. Trial results can then be found more efficiently using TRNs to search publication databases [@huser2013po]. Publications can then also be automatically linked within trial registrations, as is currently done with ClinicalTrials.gov using TRN metadata from PubMed [@nationallibraryofmedicine2017c]. Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings. <!--# add policy background on why pub in reg --> Results references may also be entered manually in many registries, with varying degrees of structure. ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas DRKS offers an unstructured free-text field. <!--# add some background on ctgov vs. drks. and how they include trns and publications. and ctgov and pubmed by same nlm so may be better at seing those ids -->While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].

Trial evidence appears to be insufficiently linked in both registrations and publications. Previous studies on trial registrations found links to results publications in as few as 13% to as many as 60% of trials [@bashir2017sr; @huser2013po; @wetering2012po]. TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b; @huser2013jamia; @wetering2012po]. The sampling strategies used in these previous studies limit our ability to draw conclusions about registration-publication linking. Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b]. Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies investigating publication links in DRKS. In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practices.

@@

The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017. We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs, which may be applied to other trial cohorts. <!--# We did not pre-specify the analyses. We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication. We also looked at reporting PMC-version full-text. -->

@@

We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset [@riedel2021a]. In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.

We <!--# programmatically -->downloaded data from both registries on 15 May 2021. <!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data. -->We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti]. As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm]. <!--# ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information. For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS. We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").-->

We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->

@@

After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021 for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.

@@

As reference data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL, when available. We used regexes to extract any DOIs and PMIDs from the citation and URL. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We generated descriptive statistics on trial and publication characteristics, overall and by registry. We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables. Regressions were performed as exploratory analyses, and all variables were included in each regression model.<!--#no model selection or fitting was conducted; the explanatory variables were tested a priori to verify there was no violation of the assumption of the linearity of the logit TODO, ask Murph--> In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata. Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs. manually linked publications in ClinicalTrials.gov. To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration. <!--#add automatic links to regression for pub in registry? in which case would need to recode NA as FALSE. To explore whether automated links explains registry differences in publication links in the registry...-->

@@

The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.

```{r flow-trials, fig.cap = "Trials and publications screening", out.width="50%", out.height="50%"}
knitr::include_graphics(here("docs", "figures", "flow-trials.png"))
```

@@

```{r tbl-trials, results = "asis"}
# tbl <- read_rds(here("docs", "figures", "tbl-trials.rds"))
# tbl %>% 
#   as_kable_extra(booktabs = TRUE)
tbl_trials %>%

  as_kable_extra()
  # as_tibble() %>%
  # apa_table(
  #   caption = "This is a table about trials",
  #   format.args = list(na_string = ""),
  #   format = "pandoc"
  # )
# modify_caption("Characteristics of included trials with results publications")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Characteristics of included trials with results publications")
```

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata. Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_n_pct_trials(n_trials_link_trn_ft_only)` trials. The inclusion of the various link types ranged from `r report_pct_num_denom_trials(n_trials_link_trn_abs)` in abstracts to `r report_pct_num_denom_trials(n_trials_link_trn_ft)` in full-text. Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry. <!--# point estimates in table should have CI, right? -->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.-->

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  modify_caption("Registration-publication links by registry") #%>% 
  # gtsummary::as_flex_table() #%>%
  # flextable::set_caption(caption = "Registration-publication links by registry")
```

```{r plot-upset-links-reg-pub-registries, fig.cap = "Registration-publication links in ClinicalTrials.gov and DRKS", out.width="80%"}
knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
```

```{r tbl-trials-regression}
tbl_trials_regression %>% 
  modify_caption("Crude and adjusted odds ratios for factors associated with publication-registration links")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
```

Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables. Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`). Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.

```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication links in ClinicalTrials.gov and DRKS. Completion year from registry."}
knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
```

>>>>

  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract. In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links. The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft")`). ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

@@

keywords          : "clinical trials, registration, reporting, meta-research, research transparency"
wordcount         : "3551"

bibliography      : ["references.bib", "r-references.bib"]
appendix          : "appendix.Rmd"
numbersections    : yes
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

always_allow_html: true

nocite: |
  @taylorfrancisgroup2020, @crossref2020c, @nationallibraryofmedicine2017c 
---

```{r rmd-setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r packages}
library(dplyr)
library(tidyr)
library(stringr)
library(glue)
library(purrr)
library(here)
library(readr)
library(fs)

library(ggplot2)
library(showtext)

library(DiagrammeR)
library(gtsummary)
library(gt)
library(papaja)
library(knitr)
library(kableExtra)
```

```{r r-citations}
r_citation <- cite_r()                              # base R
r_refs("r-references.bib", append = FALSE)          # references to loaded packages
create_bib(
  c("loggit", "rentrez", "roadoi"),
  "r-references.bib", append = TRUE) # references to additional packages
```

```{r figures-setup}
# figures (plots and tables)

# Prepare directory
dir_figures <- dir_create(here("manuscript", "figures"))

# theme_gtsummary_compact()

# Set themes
lightgray <- "gray75" #"coral3" # #"gray60"
darkgray <- "gray35" #"gray10"
registry_colors <- c("DRKS" = darkgray, "ClinicalTrials.gov" = lightgray)
# registry_linetype <- c("DRKS" = "longdash", "ClinicalTrials.gov" = "solid")

# font_add_google("Poppins", "Poppins")
font_add_google("Roboto", "Roboto")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()

theme_set(theme_light(
  # base_size = 18,
  base_family = "Roboto"
))

theme_update(
  text = element_text(family = "Roboto"),
  # axis.title = element_text(size = 16),
  axis.text = element_text(family = "Roboto Mono",
                           size = 8),
)
```

```{r raw-data, cache = TRUE}
# Dataset created in https://github.com/maia-sh/intovalue-data

# TODO: create a github release and change link to version; for now using commit link: https://docs.github.com/en/github/administering-a-repository/releasing-projects-on-github/managing-releases-in-a-repository
# "https://github.com/maia-sh/intovalue-data/raw/main/data/processed/trials.csv?raw=true"

dir <- dir_create(here("data"))

# Trials dataset
trials_raw <- 
  read_csv("https://github.com/maia-sh/intovalue-data/blob/b7909d4059e1435695e6e0be56d4da468b1f5a96/data/processed/trials.csv?raw=true") %>% 
  write_csv(fs::path(dir, "trials-raw.csv"))

# PubMed metadata
pubmed_si <- read_rds("https://github.com/maia-sh/intovalue-data/blob/a95fd7e93f1838a595e747f7f756a13a49fb66bd/data/processed/pubmed/pubmed-si.rds?raw=true") %>% 
  write_rds(fs::path(dir, "pubmed-si.rds"))

# Query dates
download.file(
  "https://github.com/maia-sh/intovalue-data/raw/6b97fb62804aae1bab6c85bd2817baa027207bab/queries.log",
  fs::path(dir, "queries.log")
)

query_logs <- loggit::read_logs(fs::path(dir, "queries.log"))

# Extract most recent query dates
get_latest_query <- function(query, logs) {
  logs %>%
    filter(log_msg == query) %>%
    arrange(desc(timestamp)) %>%
    slice_head(n = 1) %>%
    pull(timestamp) %>%
    as.Date.character() %>%
    format("%d %B %Y")
}

pubmed_query_date <- get_latest_query("PubMed", query_logs)
aact_query_date <- get_latest_query("AACT", query_logs)
drks_query_date <- get_latest_query("DRKS", query_logs)
```

```{r analysis, cache = TRUE}
dir_ls(here("analysis")) %>%
  walk(source)
```

```{r abstract-vars, cache = TRUE, dependson = "analysis"}
abs_n_trials_link_full <- 
  report_trials(n_trials_link_full, pattern = "{p}% ({n}/{N})")
abs_n_trials_link_ft_abs <- 
  report_trials(n_trials_link_ft_abs, pattern = "{p}% ({n})")
abs_n_trials_link_none <- 
  report_trials(n_trials_link_none)
abs_n_trials_link_ft <- 
  inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{n}, {p}% of trials")
abs_n_trials_link_reg_ctgov_auto <- 
  report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")
```

@@

Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses. The validity of health evidence synthesis is jeopardized if relevant trial data are either left out due to an inability to find results through automated means based on inadequate trial metadata, or duplicated due to an inability to connect publications and their registrations. Incomplete or duplicated evidence threatens the validity of clinical decision-making and potentially misinforms decisions regarding the launch of subsequent clinical trials, which may affect the wellbeing of patients enrolled in them [@chan2014tl; @councilforinternationalorganizationsofmedicalsciences2016]<!--# undermines the justification for insurance reimbursements, drives the misallocation of healthcare resources-->. Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b]. Linking trial registrations and results publications makes each more findable, reducing the risk of information loss or duplication and increasing clinical research value [@kubben2019]<!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->. Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].

@@

The inclusion of TRNs in publication and bibliographic metadata especially enhances machine readability and discoverability. Trial results can then be found more efficiently using TRNs to search publication databases [@huser2013po]. Publications can then also be automatically linked within trial registrations, as is currently done with ClinicalTrials.gov using TRN metadata from PubMed [@nationallibraryofmedicine2017c]. Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings. <!--# add policy background on why pub in reg --> Results references may also be entered manually in many registries, with varying degrees of structure. ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas the German Clinical Trials Registry (DRKS) offers an unstructured free-text field. <!--# add some background on ctgov vs. drks. and how they include trns and publications. and ctgov and pubmed by same nlm so may be better at seing those ids -->While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].

Trial evidence appears to be insufficiently linked in both registrations and publications. Previous studies on trial registrations found links to results publications in as few as 13% to as many as 60% of trials [@bashir2017sr; @huser2013po; @wetering2012po]. TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b; @huser2013jamia; @wetering2012po]. The sampling strategies used in previous studies limit our ability to draw conclusions about registration-publication linking. Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b]. Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies investigating publication links in DRKS. In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practices.

@@

The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017. We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs, which may be applied to other trial cohorts. <!--# We did not pre-specify the analyses. We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication. We also looked at reporting PMC-version full-text. -->

@@

We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021c and referred to as the "IntoValue" dataset [@riedel2021a]. The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.

We <!--# programmatically -->downloaded data from both registries on `r ifelse(aact_query_date == drks_query_date, aact_query_date, stop())`. <!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data. -->We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti]. As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm]. <!--# ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information. For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS. We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").-->

We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->

@@

After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021c for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.

@@

As publication data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL, when available. We used regexes to extract any DOIs and PMIDs from the citation and URL. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We generated descriptive statistics on trial and publication characteristics, overall and by registry. We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables. Regressions were performed as exploratory analyses, and all variables were included in each regression model. There was no model selection or fitting, or correction for multiple testing. In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata. Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs. manually linked publications in ClinicalTrials.gov. To explore registry differences for only manually linked publications, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration. <!--#add automatic links to regression for pub in registry? in which case would need to recode NA as FALSE. To explore whether automated links explains registry differences in publication links in the registry...-->

@@

The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Appendix \@ref(app-flow-trials) provides a flow diagram of the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.

@@

```{r tbl-trials, results = "asis"}
as_kable(tbl_trials, format = 'pipe')
```

Per our inclusion criteria all trials in our sample (n = `r n_trials_ft`) were registered and had a publication, however only `r report_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning the publication linked in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata. Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_trials(n_trials_link_ft_abs_reg)`. An additional `r report_trials(n_trials_link_ft_abs - n_trials_link_ft_abs_reg)` met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract. In contrast, we found `r report_trials(n_trials_link_none)` trials with no links in either the registration or the publication. The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_trials(n_trials_link_ft_only)` trials. The inclusion of the various link types ranged from `r inline_text(tbl_link_by_reg, has_iv_trn_abstract, stat_0, pattern = "{p}% ({n}/{N})")` in abstracts to `r inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{p}% ({n}/{N})")` in full-text. Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry. <!--# point estimates in table should have CI, right? -->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.-->

```{r tbl-link-by-reg}
as_kable(tbl_link_by_reg, format = "pipe")
```

```{r plot-upset-links-reg-pub-registries, fig.cap = "Trials with all combinations of links between registration and publication in ClinicalTrials.gov and DRKS", out.width="80%"}
knitr::include_graphics(path(dir_figures, "plot-upset-link-type.pdf"))
```

```{r tbl-trials-regression}
# options(knitr.kable.NA = "**")
as_kable(tbl_trials_regression, format = "pipe") #%>% 
  # kable_styling(latex_options = "scale_down")
  # landscape()
```

Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables. Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`). Figure \@ref(fig:plot-link-year) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.

```{r plot-link-year, fig.cap = "Percentage of trials with linked registrations and publications by trial completion year in ClinicalTrials.gov and DRKS. Completion year from registry.", out.width = "95%"}
knitr::include_graphics(path(dir_figures, "plot-link-year.pdf"))
```

Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with link type were generally more likely to have another type of link. <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`). Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`). Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_1", pattern = "{p}%")` vs. `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`). Appendix \@ref(app-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
